# Basic R for Graphics"

###  Scott Karr
###  HW4 Test Score Classification
###  11.22.2015
### Packages:  ggplot2, data.tables

Perfect Colinearity:
  see video [https://www.youtube.com/watch?v=DDRQYKVFoP0]
  
Multilinearity:
  see video [https://www.youtube.com/watch?v=O4jDva9B3fw]

NYC Dept of Ed. standardized test score data Grades 3-8:  
  see:  [http://schools.nyc.gov/Accountability/data/TestResults/ELAandMathTestResults]

# Setup and Load data
1. Load the New York City English Language Arts test scores dataset by District, Grade and Year. ggplot2 package.  Identify variables as numeric or categorical (factors).

```{r}
## Load Data Frame from website
require(data.table)
require(ggplot2)

#ELAByDistrictUrl <- "/Users/scottkarr/Downloads/District ELA Results 2013-2015.csv"
ELAByDistrictUrl <- "https://raw.githubusercontent.com/scottkarr/hw4/master/District%20ELA%20Results%202013-2015.csv"
df_ELAByDistrict <- read.table(file = ELAByDistrictUrl, header = TRUE, sep = ",")
dt_ELAByDistrict <- data.table(df_ELAByDistrict)
```

## Generate Descriptive Statistics
2. Generate summary level descriptive statistics:

```{r}
head(df_ELAByDistrict)
str(df_ELAByDistrict)
```

## Prepare Dataset for Graphics
3. Index by District

```{r}
## Create column class as key column for indexing
setkey(dt_ELAByDistrict, District)
```

4. Filter on "2015" and limit dataset to percentages by level and District, Grade, Year.
```{r}
ans2 <- dt_ELAByDistrict[ Year == 2015 , .(District, Grade, Year, Lvl1Pct, Lvl2Pct, Lvl3Pct, Lvl4Pct, PassPct)]
```

5.  Histogram showing Passing percentages for 2015 by District.  Note 4 distinct groupings which are related to levels
```{r}
ggplot(data = ans2) + geom_histogram(aes(x = PassPct))
```

6. Scatterplot superimposing scatterplot and box diagram, grouped by Grade.  Data compares 2015 ELA level1 percentages to passing percentages(level3 & level4).  Note the negative correlation since passing and level 1 scores are mutually exclusive, one would expect an inverse relationship between passing and level1 1 percentages.
```{r}
g <- ggplot(ans2, aes(x = PassPct, y = Lvl1Pct)) + geom_point() + geom_boxplot()
g <- g + geom_point(aes(color = Grade)) + facet_wrap(~Grade)
g
```

7 a. Scatterplot superimposing scatterplot and box diagram, grouped by Grade.  Data compares 2015 ELA level2 percentages to passing percentages(level3 & level4).  Note the lack of correlation between passing and level2 percentages.  Passing and level2 scores are still mutually exclusive but their percentages aren't correlated.  Note however that for Grades that have higher percentages of level2 scores they have lower percentages of level3 scores.  (Compare to 7.)

7 b. Grades 3 & 8 appear to have lower scores (higher percentage passing correlate to level2 than level3 scores) perhaps attributable to less experience with the newer tests.  The graphics help identify this pattern which can be explored further

```{r}
g <- ggplot(ans2, aes(x = PassPct, y = Lvl2Pct)) + geom_point() + geom_boxplot()
g <- g + geom_point(aes(color = Grade)) + facet_wrap(~Grade)
g
```

8. Scatterplot superimposing scatterplot and box diagram, grouped by Grade.  Data compares 2015 ELA level3 percentages to passing percentages(level3 & level4).  Note the positive correlation since passing and level 3 & 4 scores are colinear.
Also note that lower level2 scores show up as higher level 3 scores.  

```{r}
g <- ggplot(ans2, aes(x = PassPct, y = Lvl3Pct)) + geom_point() + geom_boxplot()
g <- g + geom_point(aes(color = Grade)) + facet_wrap(~Grade)
g
```

9. Scatterplot superimposing scatterplot and box diagram, grouped by Grade.  Data compares 2015 ELA level4 percentages to passing percentages(level3 & level4).  Note the positive correlation between passing and level4 percentages.
```{r}
g <- ggplot(ans2, aes(x = PassPct, y = Lvl4Pct)) + geom_point() + geom_boxplot()
g <- g + geom_point(aes(color = Grade)) + facet_wrap(~Grade)
g
```

10. Scatterplot superimposing scatterplot and box diagram, grouped by Grade.  Data compares 2015 ELA level4 percentages to passing percentages(level3 & level4).  Note the positive correlation between passing and level4 percentages.

```{r}
g <- ggplot(ans2, aes(x = PassPct, y = Lvl4Pct)) + geom_point() + geom_boxplot()
g <- g + geom_point(aes(color = District)) + facet_wrap(~Grade)
g
```